---
title: "sudden gains"
author: "ryan schubert"
date: "11/18/2021"
output: html_document
---

```{r setup, include=FALSE}
library(suddengains)
library(dplyr)
library(tidyr)
"%&%" = function(a,b) paste0(a,b)
set.seed(1234)
```

## R Markdown


```{r cars}
out.dir<-"/Users/ryanschubert/Dropbox/Ryan's stuff/rush/sudden gains/"
dir<-"/Users/ryanschubert/Dropbox/Ryan's stuff/rush/ML project/datasets/3 Week IOP/"
# list.files(dir)
rawData<-fread(dir %&% "IOP_DATA_SET_6-12-2020 with Manual Corrections copy.csv" ,na=c("-"))
# colnames(rawData)[grepl("PCL_DAY",colnames(rawData))]
PCL_Data<-rawData %>% select(one_of("PAT_ID","PCL_SCORE","PCL_DAY" %&% 1:14)) 

duplicated_PAT_ID<-PCL_Data$PAT_ID[duplicated(PCL_Data$PAT_ID)] %>% unique()

PCL_Data<-PCL_Data %>% filter(!(PAT_ID %in% duplicated_PAT_ID))
```

```{r}

cutoff<-define_crit1_cutoff(data_sd=PCL_Data$PCL_SCORE,reliability=0.84)

vars<-colnames(PCL_Data)[-1]
sg_values<-identify_sg(PCL_Data,
            id_var_name = "PAT_ID",
            sg_var_list=vars,
            sg_crit1_cutoff=cutoff$standard_error_difference,
            sg_crit3 = T)

sg<-sg_values %>% select(contains("sg_"))
sg_vec<-rowSums(sg,na.rm=T) 
table(sg_vec)/length(sg_vec)

table(sg_vec)

```

so we have the transitions where a sudden gain is present, we need to identify the magnitude of these sudden gains to calculate the magnitude/threshold of a 50% reversal. We then just need to check if any subsequent time points are above the threshold. Thankfully the sudden gains package makes this easy for us and identifies if a person had a subsequent reversal using the 50% reversal threshold.



```{r}
byperson <- create_byperson(data = PCL_Data,
                    sg_crit1_cutoff=cutoff$standard_error_difference,
                    id_var_name = "PAT_ID",
                    tx_start_var_name = "PCL_SCORE",
                    tx_end_var_name = "PCL_DAY14",
                    sg_var_list=vars,
                    identify_sg_1to2=T,
                    sg_measure_name = "PCL",
                    identify = "sg")
table(byperson$sg_reversal)
fwrite(byperson,out.dir %&% "sudden_gains.csv")
```

of the 205 people who experienced at least one sudden gain, only 16 had a reversal, leaving 189 as having a sudden gain.

```{r}

sudden_gainers<-byperson %>% filter(sg_reversal==0)
```

what about in a complete case sense

```{r}
PCL_Data<-PCL_Data[complete.cases(PCL_Data),]
byperson <- create_byperson(data = PCL_Data,
                    sg_crit1_cutoff=cutoff$standard_error_difference,
                    id_var_name = "PAT_ID",
                    tx_start_var_name = "PCL_SCORE",
                    tx_end_var_name = "PCL_DAY14",
                    sg_var_list=vars,
                    identify_sg_1to2=T,
                    sg_measure_name = "PCL",
                    identify = "sg")
table(byperson$sg_reversal)
```

similar results looking at a complete case analysis. 168/390 individuals have a sudden gain, with only 12 of those having a reversal, meaning 156 individuals have a relable sudden gain

```{r}

fwrite(byperson,out.dir %&% "complete_cases_sudden_gains.csv")
```




