---
title: "compare sudden gainers"
author: "ryan schubert"
date: "12/28/2021"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(data.table)
library(plotrix)
library(ggplot2)
library(viridis)
library(lme4)
library(lmerTest)
"%&%" = function(a,b) paste0(a,b)
set.seed(1234)
```

## R Markdown

```{r cars}
pcl.dir<-"/Users/ryanschubert/Dropbox/Ryan's stuff/rush/ML project/datasets/3 Week IOP/"
data.dir<-"/Users/ryanschubert/Dropbox/Ryan's stuff/rush/sudden gains/"
out.dir<-"/Users/ryanschubert/Dropbox/Ryan's stuff/rush/sudden gains/"
```

## Including Plots

```{r}
rawData<-fread(pcl.dir %&% "IOP_DATA_SET_6-12-2020 with Manual Corrections copy.csv" ,na=c("-"))
colnames(rawData)[grepl("mst",colnames(rawData),ignore.case=T)]
PCL_Data<-rawData %>% 
  dplyr::select(one_of("PAT_ID","PCL_SCORE","PCL_DAY" %&% 1:14),gender,AGE_AT_COHORT_START,MSTYN) %>%
  mutate(gender=if_else(gender=="Female",1,0))


duplicated_PAT_ID<-PCL_Data$PAT_ID[duplicated(PCL_Data$PAT_ID)] %>% unique()

PCL_Data<-PCL_Data %>% 
  filter(!(PAT_ID %in% duplicated_PAT_ID)) %>%
  mutate(delta_PCL=PCL_SCORE-PCL_DAY14,
         CMI=delta_PCL>=10)
```

```{r}
sg<-fread(out.dir %&% "complete_cases_sudden_gains.csv")
hist(sg$sg_session_n)
sg<-sg %>% mutate(reliable_sg=case_when(
  sg_reversal==0 ~ 1,
  is.na(sg_reversal) ~ 0,
  T ~ 0)) %>% 
  dplyr::select(PAT_ID,reliable_sg)
sg_Data<-filter(PCL_Data,PAT_ID %in% sg$PAT_ID)  %>%
  inner_join(sg,by="PAT_ID")


```

so what I want to know is two things,

how do sudden gains compare to no sudden gains

how do sudden gains compare to no sudden gains conditional on each individual being a responder

First lets simply use regression based methods

# sudden gains vs no sudden gains

```{r}
modeling_data<-sg_Data %>%
  pivot_longer(PCL_SCORE:PCL_DAY14) %>%
  mutate(name=factor(name,levels=c("PCL_SCORE","PCL_DAY2","PCL_DAY3","PCL_DAY5","PCL_DAY6","PCL_DAY8","PCL_DAY10", "PCL_DAY11","PCL_DAY13","PCL_DAY14")))

mixed_model<-lmer(value ~ name*reliable_sg + AGE_AT_COHORT_START + gender +  MSTYN + (1|PAT_ID),data=modeling_data)
anova(mixed_model)

plot_data<-modeling_data %>% 
  group_by(name,reliable_sg) %>%
  summarise(mean_score=mean(value),
            ste_score=std.error(value)) %>%
  mutate(name=factor(name,levels=c("PCL_SCORE","PCL_DAY2","PCL_DAY3","PCL_DAY5","PCL_DAY6","PCL_DAY8","PCL_DAY10", "PCL_DAY11","PCL_DAY13","PCL_DAY14")))

ggplot(plot_data,aes(x=name,y=mean_score)) +
  geom_line(aes(group=reliable_sg,colour=as.factor(reliable_sg))) +
  geom_errorbar(aes(ymin=mean_score-ste_score,ymax=mean_score+ste_score,colour=as.factor(reliable_sg))) +
  geom_point(aes(shape=as.factor(reliable_sg))) +
  theme_bw()
```

#sudden gains vs no sudden gains conditional on everyone being a 10pt responder

```{r}
modeling_data<-sg_Data %>%
  filter(CMI==T) %>%
  pivot_longer(PCL_SCORE:PCL_DAY14) %>%
  mutate(name=factor(name,levels=c("PCL_SCORE","PCL_DAY2","PCL_DAY3","PCL_DAY5","PCL_DAY6","PCL_DAY8","PCL_DAY10", "PCL_DAY11","PCL_DAY13","PCL_DAY14")))

mixed_model<-lmer(value ~ name*reliable_sg + AGE_AT_COHORT_START + gender +  MSTYN + (1|PAT_ID),data=modeling_data)
anova(mixed_model)

plot_data<-modeling_data %>% 
  group_by(name,reliable_sg) %>%
  summarise(mean_score=mean(value),
            ste_score=std.error(value)) %>%
  mutate(name=factor(name,levels=c("PCL_SCORE","PCL_DAY2","PCL_DAY3","PCL_DAY5","PCL_DAY6","PCL_DAY8","PCL_DAY10", "PCL_DAY11","PCL_DAY13","PCL_DAY14")))

ggplot(plot_data,aes(x=name,y=mean_score)) +
  geom_line(aes(group=reliable_sg,colour=as.factor(reliable_sg))) +
  geom_errorbar(aes(ymin=mean_score-ste_score,ymax=mean_score+ste_score,colour=as.factor(reliable_sg))) +
  geom_point(aes(shape=as.factor(reliable_sg))) +
  theme_bw()
```

Nikki wants to do a matched pairs design


```{r}
library(MatchIt)
matched_data<-matchit(reliable_sg ~ PCL_SCORE + AGE_AT_COHORT_START + gender + MSTYN,
                      data=sg_Data,
                      distance="glm",
                      link="logit",
                      mahvars=~ PCL_SCORE + AGE_AT_COHORT_START + gender + MSTYN,
                      calipers=0.25,
                      replace=F)
```


```{r}
pairs<-matched_data$match.matrix %>% 
  as.data.frame() %>% 
  rownames_to_column() %>%
  rename(sg="rowname",no_sg="V1")
matched_subset<-sg_Data %>%
  rownames_to_column() %>%
  mutate(Match_id=case_when(
    rowname %in% pairs$sg ~ "sg",
    rowname %in% pairs$no_sg ~ "no_sg"
  )) %>%
  filter(!is.na(Match_id))
```

for each pair we need to extract the index of the sessions surrounding the sg. Then we need to extract the PCL scores of the non sg individual in the pair and extract them in a matching format as outout by the describe_sg function

```{r}
row_to_PAT_ID<-sg_Data %>% dplyr::select(PAT_ID) %>% rownames_to_column()
pairs<-pairs %>% 
  inner_join(row_to_PAT_ID,by=c("sg"="rowname")) %>%
  dplyr::select(-sg) %>%
  rename(sg="PAT_ID") %>%
  inner_join(row_to_PAT_ID,by=c("no_sg"="rowname")) %>%
  dplyr::select(-no_sg) %>%
  rename(nosg="PAT_ID")
matched_only<-sg %>% filter(PAT_ID %in% pairs$sg | PAT_ID %in% pairs$nosg)

for (i in 1:nrow(pairs)){
  sg_instance<-filter(sg_only,PAT_ID==pairs$sg[i])
  session_n<-sg_instance$sg_session_n
  indices<-session_n + c(-2,-1,0,1,2,3)
  indices<-indices[indices>=0 & indices <=9]
  indices<-indices + 17
  if(session_n==1){
    placing_indices<-c(7:11)
  }  else if (session_n==7) {
    placing_indices<-c(6:10)
  } else {
    placing_indices<-c(6:11)
  }
  matched_only[matched_only$PAT_ID == pairs$nosg[i],placing_indices]<-matched_only[matched_only$PAT_ID == pairs$nosg[i],..indices]
}

matched_only<-matched_only %>%
  mutate(Match_ID=case_when(sg_reversal==0~"sg",
          is.na(sg_reversal)~"no_sg",
          T~"no_sg"))
```

```{r}
bysg<-create_bysg(matched_only,
            sg_crit1_cutoff=6.8,
            sg_measure_name="PCL",
            id_var_name="PAT_ID",
            tx_start_var_name="PCL_SCORE",
            tx_end_var_name="PCL_DAY14",
            sg_var_list=c("PCL_SCORE","PCL_DAY2","PCL_DAY3","PCL_DAY5","PCL_DAY6","PCL_DAY8","PCL_DAY10","PCL_DAY11","PCL_DAY13","PCL_DAY14"))

sudden_gains_plot<-plot_sg(data=matched_only,
        id_var_name="PAT_ID",
        tx_start_var_name="PCL_SCORE",
        tx_end_var_name="PCL_DAY14",
        sg_pre_post_var_list=c("sg_PCL_2n","sg_PCL_1n","sg_PCL_n","sg_PCL_n1","sg_PCL_n2","sg_PCL_n3"),
        ylab = "PCL-5 Score", xlab = "Session",
        colour_group = "ggplot",
        group_var_name = "Match_ID",
        group_labels=c("no_sg","sg"),
        group_levels=c("no_sg","sg")
        )
ggsave(out.dir %&% "Sudden_gains_matched_data.png",dpi=300,height=4.9,width=7,plot=sudden_gains_plot)
```

```{r}
sg_only_outcome<-sg_Data %>% inner_join(pairs,by=c("PAT_ID"="sg")) %>% dplyr::select(PAT_ID,delta_PCL,nosg)
nosg_only_outcome<-sg_Data %>% inner_join(pairs,by=c("PAT_ID"="nosg")) %>% dplyr::select(PAT_ID,delta_PCL,sg)
comparable<-inner_join(sg_only_outcome,nosg_only_outcome,by=c("nosg"="PAT_ID"))
t.test(comparable$delta_PCL.x,comparable$delta_PCL.y,paired = T)
```




