---
title: "two sample ttest compare our data to wiedemann and sloan"
author: "ryan schubert"
date: "4/4/2022"
output: html_document
---

```{r setup, include=FALSE}
"%&%" = function(a,b) paste0(a,b)
```

## R Markdown


Sloan - pcl
In their WET sample: n=13/63, mean= 22.23, SD = 6.1, 
In their CPT sample: n=11/63, mean= 20.73, SD = 4.9


Wiedemann uses PDS, which is out of a total of 51.
Wiedemann Sample 1: n= 76/248 mean= 12.30, SD= 4.44
Wiedemann Sample 2: n= 87/234 mean= 12.11 SD= 3.83

3 week sample
n=146/465,mean=21.23288,sd=6.693323

2 week sample
n=44/235 mean=21.3617 sd=5.027859


# 3week vs sloan
```{r cars}
# 3 week vs WET
mu_3W=21.23288
n_3W=146
sd_3W=6.693323

mu_SWET=22.23
n_SWET=13
sd_SWET=6.1

mu_SCPT=20.73
n_SCPT=11
sd_SCPT=4.9

delta_mu=mu_3W-mu_SWET
s1<-(sd_3W^2)/n_3W
s2<-(sd_SWET^2)/n_SWET
denom=sqrt(s1+s2)
(t=delta_mu/denom)
df_num=(s1+s2)^2
df_den=(sd_3W^4)/((n_3W^2)*(n_3W-1)) + (sd_SWET^4)/((n_SWET^2)*(n_SWET-1))
df1<-df_num/df_den
2*pt(t,df1,lower.tail=T)

delta_mu=mu_3W-mu_SCPT
s1<-(sd_3W^2)/n_3W
s2<-(sd_SCPT^2)/n_SCPT
denom=sqrt(s1+s2)
(t=delta_mu/denom)
df_num=(s1+s2)^2
df_den=(sd_3W^4)/((n_3W^2)*(n_3W-1)) + (sd_SCPT^4)/((n_SCPT^2)*(n_SCPT-1))
df1<-df_num/df_den
2*pt(t,df1,lower.tail=F)
```

3 week not significantly different from wet or CPT in sloan

#2 week vs slaon
```{r}
mu_2W=21.3617
n_2W=44
sd_2W=5.027859

delta_mu=mu_2W-mu_SWET
s1<-(sd_2W^2)/n_2W
s2<-(sd_SWET^2)/n_SWET
denom=sqrt(s1+s2)
(t=delta_mu/denom)
df_num=(s1+s2)^2
df_den=(sd_2W^4)/((n_2W^2)*(n_2W-1)) + (sd_SWET^4)/((n_SWET^2)*(n_SWET-1))
df1<-df_num/df_den
2*pt(t,df1,lower.tail=T)

delta_mu=mu_2W-mu_SCPT
s1<-(sd_2W^2)/n_2W
s2<-(sd_SCPT^2)/n_SCPT
denom=sqrt(s1+s2)
(t=delta_mu/denom)
df_num=(s1+s2)^2
df_den=(sd_2W^4)/((n_2W^2)*(n_2W-1)) + (sd_SCPT^4)/((n_SCPT^2)*(n_SCPT-1))
df1<-df_num/df_den
2*pt(t,df1,lower.tail=F)

```
2week not signif different from wet or cpt in sloan



# 3 week vs wiedemann
```{r}
mu_3W=21.23288/80
n_3W=146
sd_3W=6.693323/80

n_WS1=76 
mu_WS1=12.30/51
sd_WS1= 4.44/51

n_WS2= 87 
mu_WS2= 12.11/51
sd_WS2= 3.83/51

delta_mu=mu_3W-mu_WS1
s1<-(sd_3W^2)/n_3W
s2<-(sd_WS1^2)/n_WS1
denom=sqrt(s1+s2)
(t=delta_mu/denom)
df_num=(s1+s2)^2
df_den=(sd_3W^4)/((n_3W^2)*(n_3W-1)) + (sd_WS1^4)/((n_WS1^2)*(n_WS1-1))
df1<-df_num/df_den
2*pt(t,df1,lower.tail=F)

delta_mu=mu_3W-mu_WS2
s1<-(sd_3W^2)/n_3W
s2<-(sd_WS2^2)/n_WS2
denom=sqrt(s1+s2)
(t=delta_mu/denom)
df_num=(s1+s2)^2
df_den=(sd_3W^4)/((n_3W^2)*(n_3W-1)) + (sd_WS2^4)/((n_WS2^2)*(n_WS2-1))
df1<-df_num/df_den
2*pt(t,df1,lower.tail=F)
```

```{r}
mu_3W=19.33161/80
n_3W=193
sd_3W=6.96598/80

n_WS1=76 
mu_WS1=12.30/51
sd_WS1= 4.44/51

n_WS2= 87 
mu_WS2= 12.11/51
sd_WS2= 3.83/51

delta_mu=mu_3W-mu_WS1
s1<-(sd_3W^2)/n_3W
s2<-(sd_WS1^2)/n_WS1
denom=sqrt(s1+s2)
(t=delta_mu/denom)
df_num=(s1+s2)^2
df_den=(sd_3W^4)/((n_3W^2)*(n_3W-1)) + (sd_WS1^4)/((n_WS1^2)*(n_WS1-1))
df1<-df_num/df_den
2*pt(t,df1,lower.tail=F)

delta_mu=mu_3W-mu_WS2
s1<-(sd_3W^2)/n_3W
s2<-(sd_WS2^2)/n_WS2
denom=sqrt(s1+s2)
(t=delta_mu/denom)
df_num=(s1+s2)^2
df_den=(sd_3W^4)/((n_3W^2)*(n_3W-1)) + (sd_WS2^4)/((n_WS2^2)*(n_WS2-1))
df1<-df_num/df_den
2*pt(t,df1,lower.tail=F)
```

```{r}
mu_2W=21.3617/80
n_2W=44
sd_2W=5.027859/80

delta_mu=mu_2W-mu_WS1
s1<-(sd_2W^2)/n_2W
s2<-(sd_WS1^2)/n_WS1
denom=sqrt(s1+s2)
(t=delta_mu/denom)
df_num=(s1+s2)^2
df_den=(sd_2W^4)/((n_2W^2)*(n_2W-1)) + (sd_WS1^4)/((n_WS1^2)*(n_WS1-1))
df1<-df_num/df_den
2*pt(t,df1,lower.tail=F)

delta_mu=mu_2W-mu_WS2
s1<-(sd_2W^2)/n_2W
s2<-(sd_WS2^2)/n_WS2
denom=sqrt(s1+s2)
(t=delta_mu/denom)
df_num=(s1+s2)^2
df_den=(sd_2W^4)/((n_2W^2)*(n_2W-1)) + (sd_WS2^4)/((n_WS2^2)*(n_WS2-1))
df1<-df_num/df_den
2*pt(t,df1,lower.tail=F)
```


```{r}
mu_2W=17.85897/80
n_2W=78
sd_2W=5.945039/80

delta_mu=mu_2W-mu_WS1
s1<-(sd_2W^2)/n_2W
s2<-(sd_WS1^2)/n_WS1
denom=sqrt(s1+s2)
(t=delta_mu/denom)
df_num=(s1+s2)^2
df_den=(sd_2W^4)/((n_2W^2)*(n_2W-1)) + (sd_WS1^4)/((n_WS1^2)*(n_WS1-1))
df1<-df_num/df_den
2*pt(t,df1,lower.tail=T)

delta_mu=mu_2W-mu_WS2
s1<-(sd_2W^2)/n_2W
s2<-(sd_WS2^2)/n_WS2
denom=sqrt(s1+s2)
(t=delta_mu/denom)
df_num=(s1+s2)^2
df_den=(sd_2W^4)/((n_2W^2)*(n_2W-1)) + (sd_WS2^4)/((n_WS2^2)*(n_WS2-1))
df1<-df_num/df_den
2*pt(t,df1,lower.tail=T)
```


# Z tests for proportion

3week vs sloan

```{r}
h_3W=146
n_3W=465

h_SWET=13
n_SWET=63

p_hat=(h_3W+h_SWET)/(n_3W+n_SWET)

p_3W=h_3W/n_3W
p_SWET=h_SWET/n_SWET

denom=sqrt(p_hat*(1-p_hat)*((1/n_SWET) + (1/n_3W)))
Z=(p_3W-p_SWET)/denom
pnorm(Z,lower.tail = F)

h_3W=146
n_3W=465

h_SCPT=11
n_SCPT=63

p_hat=(h_3W+h_SCPT)/(n_3W+n_SCPT)

p_3W=h_3W/n_3W
p_SCPT=h_SCPT/n_SCPT

denom=sqrt(p_hat*(1-p_hat)*((1/n_SCPT) + (1/n_3W)))
Z=(p_3W-p_SCPT)/denom
pnorm(Z,lower.tail = F)
```

2week vs sloan

```{r}
h_2W=44
n_2W=235

h_SWET=13
n_SWET=63

p_hat=(h_2W+h_SWET)/(n_2W+n_SWET)

p_2W=h_2W/n_2W
p_SWET=h_SWET/n_SWET

denom=sqrt(p_hat*(1-p_hat)*((1/n_SWET) + (1/n_2W)))
Z=(p_2W-p_SWET)/denom
pnorm(Z,lower.tail = F)

h_2W=44
n_2W=235

h_SCPT=11
n_SCPT=63

p_hat=(h_2W+h_SCPT)/(n_2W+n_SCPT)

p_2W=h_2W/n_2W
p_SCPT=h_SCPT/n_SCPT

denom=sqrt(p_hat*(1-p_hat)*((1/n_SCPT) + (1/n_2W)))
Z=(p_2W-p_SCPT)/denom
pnorm(Z,lower.tail = F)
```


3week vs wiedemann

```{R}
h_3W=146
n_3W=465

h_WS1=76
n_WS1=248

p_hat=(h_3W+h_WS1)/(n_3W+n_WS1)

p_3W=h_3W/n_3W
p_WS1=h_WS1/n_WS1

denom=sqrt(p_hat*(1-p_hat)*((1/n_WS1) + (1/n_3W)))
Z=(p_3W-p_WS1)/denom
pnorm(Z,lower.tail = F)

h_3W=146
n_3W=465

h_WS2=87
n_WS2=234

p_hat=(h_3W+h_WS2)/(n_3W+n_WS2)

p_3W=h_3W/n_3W
p_WS2=h_WS2/n_WS2

denom=sqrt(p_hat*(1-p_hat)*((1/n_WS2) + (1/n_3W)))
Z=(p_3W-p_WS2)/denom
pnorm(Z,lower.tail = T)

```

2week vs wiedemann

```{R}
h_2W=44
n_2W=235

h_WS1=76
n_WS1=248

p_hat=(h_2W+h_WS1)/(n_2W+n_WS1)

p_2W=h_2W/n_2W
p_WS1=h_WS1/n_WS1

denom=sqrt(p_hat*(1-p_hat)*((1/n_WS1) + (1/n_2W)))
Z=(p_2W-p_WS1)/denom
pnorm(Z,lower.tail = T)

h_2W=44
n_2W=235

h_WS2=87
n_WS2=234

p_hat=(h_2W+h_WS2)/(n_2W+n_WS2)

p_2W=h_2W/n_2W
p_WS2=h_WS2/n_WS2

denom=sqrt(p_hat*(1-p_hat)*((1/n_WS2) + (1/n_2W)))
Z=(p_2W-p_WS2)/denom
pnorm(Z,lower.tail = T)

```

