---
title: "3 month available compare dropout data"
author: "ryan schubert"
date: "5/11/2022"
output: html_document
---

```{r setup, include=FALSE}
library(data.table)
library(dplyr)
library(psych)
library(readxl)
"%&%" = function(a,b) paste0(a,b)
checkCol=function(string,df,ignore.case=T){
  colnames(df)[grepl(string,colnames(df),ignore.case=ignore.case)]
}
```

We just want to check if people who have 3 month followup data differ in any meaningful way to people who do not have 3 month followup in both ITPs

do they differ in terms of
PCL-5, PHQ-9, PTCI

number of sudden gains

and gender, age, race, mstyn, ethnicity



```{r}
sg.dir<-"/Users/ryanschubert/Dropbox/Ryan's stuff/rush/sudden gains/"
three_week_mrns<-fread(sg.dir %&% "three_week_sufficient_data.csv") %>% select(MRN) %>% unlist()
dir<-"/Users/ryanschubert/Dropbox/Ryan's stuff/rush/cannabis use/"
completeData<-read_xlsx(dir %&% 'IOP_DATA_SET_6-12-2020 with Manual Corrections.xlsx',na = "-") %>%
  select(one_of(c("MRN","PAT_ID","COHORT","COHORT_TYPE","AGE_AT_COHORT_START","FirstRace","gender","Ethnicity","DAYSABST" %&% 1:8,"DAYSABST" %&% 1:8 %&% "_POST","PCL_SCORE","PCL_SCORE_3MON","PCL" %&% 1:20,"PHQ" %&% 1:9,"PCL_DAY" %&% 2:14,"PHQ_SCORE","PHQ_DAY" %&% 2:14,"PHQ_SCORE_POST","AUDITC" %&% 1:3,"PTCI" %&% 1:33,"NSI" %&% 1:22,"MSTYN","PTCI_DAY4","PTCI_DAY9","PTCI_DAY14")),contains("PTCI_SCORE")) %>% 
  filter(MRN %in% three_week_mrns) %>%
  mutate(followup_dropout=is.na(PCL_SCORE_3MON),
         FirstRace=if_else(FirstRace=="White","White","Non-White"),
         Ethnicity=if_else(Ethnicity=="Hispanic or Latino","Hispanic or Latino","Not Hispanic or Latino"),
         alcohol_use=DAYSABST1 + DAYSABST1_POST,
         cannabis_use=DAYSABST2 + DAYSABST2_POST,
         stimulant_use=DAYSABST3 + DAYSABST3_POST,
         heroine_use=DAYSABST4 + DAYSABST4_POST,
         opiate_use=DAYSABST5 + DAYSABST5_POST,
         tobacco_use=DAYSABST6 + DAYSABST6_POST,
         steroid_use=DAYSABST7 + DAYSABST7_POST,
         other_use=DAYSABST8 + DAYSABST8_POST)
```


lets test the three scales we care about longitudinally

PCL-5

```{r}

pcl5<-completeData %>%
  select(MRN,MSTYN,FirstRace,gender,Ethnicity,followup_dropout,PCL_SCORE,contains("PCL_DAY")) %>%
  pivot_longer(contains("PCL")) %>%
  mutate(day=case_when(
    grepl("PCL_SCORE",name) ~ "0",
    T ~ gsub("PCL_DAY","",name) 
  ),
  day=as.numeric(day))
  
#plot
pcl5_plot<-pcl5 %>%
  group_by(day,followup_dropout) %>%
  summarise(mean_score=mean(value,na.rm=T),
            se_score=std.error(value,na.rm=T))
ggplot(pcl5_plot) +
  geom_path(aes(x=day,y=mean_score,colour=followup_dropout)) +
  geom_errorbar(aes(x=day,y=mean_score,ymax=mean_score+se_score,ymin=mean_score-se_score,colour=followup_dropout)) +
  geom_point(aes(x=day,y=mean_score,shape=followup_dropout))

m1<-lmer(value ~ day*followup_dropout + MSTYN + FirstRace + gender + Ethnicity + (1 | MRN),data=pcl5)
m2<-lmer(value ~ day*followup_dropout + MSTYN + FirstRace + gender + Ethnicity + (day | MRN),data=pcl5)
summary(m2)
#test
```

no significant interaction between followup dropout and time, no significant effect of followup dropout on pcl5

phq9

```{r}
phq9<-completeData %>%
  select(MRN,MSTYN,FirstRace,gender,Ethnicity,followup_dropout,PHQ_SCORE,contains("PHQ_DAY")) %>%
  pivot_longer(contains("PHQ")) %>%
  mutate(day=case_when(
    grepl("PHQ_SCORE",name) ~ "0",
    T ~ gsub("PHQ_DAY","",name) 
  ),
  day=as.numeric(day))
  
#plot
phq9_plot<-phq9 %>%
  group_by(day,followup_dropout) %>%
  summarise(mean_score=mean(value,na.rm=T),
            se_score=std.error(value,na.rm=T))
ggplot(phq9_plot) +
  geom_path(aes(x=day,y=mean_score,colour=followup_dropout)) +
  geom_errorbar(aes(x=day,y=mean_score,ymax=mean_score+se_score,ymin=mean_score-se_score,colour=followup_dropout)) +
  geom_point(aes(x=day,y=mean_score,shape=followup_dropout))

m1<-lmer(value ~ day*followup_dropout + MSTYN + FirstRace + gender + Ethnicity + (1 | MRN),data=phq9)
m2<-lmer(value ~ day*followup_dropout + MSTYN + FirstRace + gender + Ethnicity + (day | MRN),data=phq9)
anova(m1,m2)
summary(m2)

```

no difference in phq9 score over time or at baseline between individuals who have dropped out at 3 month and those who did not

```{r}
ptci<-completeData %>%
  select(MRN,MSTYN,FirstRace,gender,Ethnicity,followup_dropout,contains("PTCI_")) %>%
  select(-PTCI_SCORE_6MON) %>%
  pivot_longer(contains("PTCI")) %>%
  mutate(day=case_when(
    grepl("PTCI_SCORE_POST",name) ~ "15",
    grepl("PTCI_SCORE",name) ~ "0",
    T ~ gsub("PTCI_DAY","",name)
  ),
  day=as.numeric(day))

ptci_plot<-ptci %>%
  group_by(day,followup_dropout) %>%
  summarise(mean_score=mean(value,na.rm=T),
            se_score=std.error(value,na.rm=T))
ggplot(ptci_plot) +
  geom_path(aes(x=day,y=mean_score,colour=followup_dropout)) +
  geom_errorbar(aes(x=day,y=mean_score,ymax=mean_score+se_score,ymin=mean_score-se_score,colour=followup_dropout)) +
  geom_point(aes(x=day,y=mean_score,shape=followup_dropout))

m1<-lmer(value ~ day*followup_dropout + MSTYN + FirstRace + gender + Ethnicity + (1 | MRN),data=ptci)
m2<-lmer(value ~ day*followup_dropout + MSTYN + FirstRace + gender + Ethnicity + (day | MRN),data=ptci)
anova(m1,m2)
summary(m2)
```
no difference in ptci score over time


and gender, age, race, mstyn, ethnicity
```{r}
chisq.test(completeData$gender,completeData$followup_dropout)
chisq.test(completeData$FirstRace,completeData$followup_dropout)
chisq.test(completeData$Ethnicity,completeData$followup_dropout)
chisq.test(completeData$MSTYN,completeData$followup_dropout)
t.test(AGE_AT_COHORT_START ~ followup_dropout,data=completeData)
```
age is the only things significantly different, however age will liekly not survive a conservitive multiple test correction such as bonferroni

magnitude and number of sudden gainers

```{r}
cutoff<-define_crit1_cutoff(data_sd=completeData$PCL_SCORE,reliability=0.84)
PCL_Data<-completeData %>% select(MRN,PCL_SCORE,contains("PCL_DAY"))
vars<-colnames(PCL_Data)[2:11]
byperson <- create_byperson(data = completeData,
                    sg_crit1_cutoff=cutoff$reliable_change_value,
                    id_var_name = "MRN",
                    tx_start_var_name = "PCL_SCORE",
                    tx_end_var_name = "PCL_DAY14",
                    sg_var_list=vars,
                    identify_sg_1to2=T,
                    sg_measure_name = "PCL",
                    identify = "sg")
byperson <- mutate(byperson,sg_stable=if_else(!is.na(sg_reversal) & sg_reversal==0,1,0))
completeData<-byperson %>% select(MRN,sg_stable) %>%
  inner_join(completeData,by=c("MRN"))
chisq.test(completeData$sg_stable,completeData$followup_dropout)
table(completeData$sg_stable,completeData$followup_dropout)


```
hm so people who had a stable sg were less likely to dropout at three month followup than people without. Wonder if this holds true in the two week data


```{r}

two_week_mrns<-fread(sg.dir %&% "two_week_sufficient_data.csv") %>% select(mrn) %>% unlist()
dir<-"/Users/ryanschubert/Dropbox/Ryan's stuff/rush/2week/"
list.files(dir)
rawData<-fread(dir %&% "Two_week_with_CAPS_2_17_22_extended.csv" ,na=c("-")) %>%
  select(mrn,contains("pcl") | contains("phq") | contains("daysabst") | contains("ptci") | contains("nsi"),"FirstRace (group)",Ethnicity,Sex,Age,cohort_type_base, contains("mst")) %>% 
  filter(mrn %in% two_week_mrns) %>%
  mutate(followup_dropout=is.na(pcl5_score_pastmonth_3_month_followup),
     FirstRace=if_else(`FirstRace (group)`=="White","White","Non-White"),
     Ethnicity=if_else(Ethnicity=="Hispanic or Latino","Hispanic or Latino","Not Hispanic or Latino"),
     mstyn = case_when(mst1_base == 1 ~ 1,
                       mst2_base == 1 ~ 1,
                       mst3_base == 1 ~ 1,
                       cohort_type_base == 2 ~ 1,
                       T ~ 0))
```


```{r}
pcl5<-rawData %>%
  select(mrn,mstyn,FirstRace,Sex,Ethnicity,followup_dropout,contains("pcl5_score")) %>%
  select(-contains("_followup")) %>%
  pivot_longer(contains("pcl5")) %>%
  mutate(day=case_when(
    name == "pcl5_score_pastmonth_base" ~ 0,
    name == "pcl5_score_monday_week_1" ~ 1,
    name == "pcl5_score_wednesday_week_1" ~ 3,
    name == "pcl5_score_friday_week_1" ~ 5,
    name == "pcl5_score_monday_week_2" ~ 6,
    name == "pcl5_score_wednesday_week_2" ~ 8,
    name == "pcl5_score_post" ~ 10),
  day=as.numeric(day))
  
#plot
pcl5_plot<-pcl5 %>%
  group_by(day,followup_dropout) %>%
  summarise(mean_score=mean(value,na.rm=T),
            se_score=std.error(value,na.rm=T))
ggplot(pcl5_plot) +
  geom_path(aes(x=day,y=mean_score,colour=followup_dropout)) +
  geom_errorbar(aes(x=day,y=mean_score,ymax=mean_score+se_score,ymin=mean_score-se_score,colour=followup_dropout)) +
  geom_point(aes(x=day,y=mean_score,shape=followup_dropout))

m1<-lmer(value ~ day*followup_dropout + mstyn + FirstRace + Sex + Ethnicity + (1 | mrn),data=pcl5)
m2<-lmer(value ~ day*followup_dropout + mstyn + FirstRace + Sex + Ethnicity + (day | mrn),data=pcl5)
anova(m1,m2)
summary(m2)

```

phq 


```{r}
phq9<-rawData %>%
  select(mrn,mstyn,FirstRace,Sex,Ethnicity,followup_dropout,contains("phq9_score")) %>%
  select(-contains("_followup")) %>%
  pivot_longer(contains("phq9")) %>%
  mutate(day=case_when(
    name == "phq9_score_base" ~ 0,
    name == "phq9_score_monday_week_1" ~ 1,
    name == "phq9_score_wednesday_week_1" ~ 3,
    name == "phq9_score_friday_week_1" ~ 5,
    name == "phq9_score_monday_week_2" ~ 6,
    name == "phq9_score_wednesday_week_2" ~ 8,
    name == "phq9_score_post" ~ 10),
  day=as.numeric(day))
  
#plot
phq9_plot<-phq9 %>%
  group_by(day,followup_dropout) %>%
  summarise(mean_score=mean(value,na.rm=T),
            se_score=std.error(value,na.rm=T))
ggplot(phq9_plot) +
  geom_path(aes(x=day,y=mean_score,colour=followup_dropout)) +
  geom_errorbar(aes(x=day,y=mean_score,ymax=mean_score+se_score,ymin=mean_score-se_score,colour=followup_dropout)) +
  geom_point(aes(x=day,y=mean_score,shape=followup_dropout))

m1<-lmer(value ~ day*followup_dropout + mstyn + FirstRace + Sex + Ethnicity + (1 | mrn),data=phq9)
m2<-lmer(value ~ day*followup_dropout + mstyn + FirstRace + Sex + Ethnicity + (day | mrn),data=phq9)
anova(m1,m2)
summary(m2)

```

```{r}

ptci<-rawData %>%
  select(mrn,mstyn,FirstRace,Sex,Ethnicity,followup_dropout,contains("ptci_score")) %>%
  select(-contains("_followup")) %>%
  pivot_longer(contains("ptci")) %>%
  mutate(day=case_when(
    name == "ptci_score_base" ~ 0,
    name == "ptci_score_thursday_week_1" ~ 4,
    name == "ptci_score_thursday_week_2" ~ 9,
    name == "ptci_score_post" ~ 10),
  day=as.numeric(day))
  
#plot
ptci_plot<-ptci %>%
  group_by(day,followup_dropout) %>%
  summarise(mean_score=mean(value,na.rm=T),
            se_score=std.error(value,na.rm=T))
ggplot(ptci_plot) +
  geom_path(aes(x=day,y=mean_score,colour=followup_dropout)) +
  geom_errorbar(aes(x=day,y=mean_score,ymax=mean_score+se_score,ymin=mean_score-se_score,colour=followup_dropout)) +
  geom_point(aes(x=day,y=mean_score,shape=followup_dropout))

m1<-lmer(value ~ day*followup_dropout + mstyn + FirstRace + Sex + Ethnicity + (1 | mrn),data=ptci)
m2<-lmer(value ~ day*followup_dropout + mstyn + FirstRace + Sex + Ethnicity + (day | mrn),data=ptci)
anova(m1,m2)
summary(m2)


```

phq and ptci are not significantly different either


and gender, age, race, mstyn, ethnicity and number of sudden gains

```{r}
chisq.test(rawData$Sex,rawData$followup_dropout)
chisq.test(rawData$mstyn,rawData$followup_dropout)
chisq.test(rawData$Ethnicity,rawData$followup_dropout)
chisq.test(rawData$FirstRace,rawData$followup_dropout)
t.test(Age ~ followup_dropout,data=rawData)
```

age didn't replicate nor did sex and mstyn. differences in sex and mst status may have more to do with changes in structure between the IOPs than anything

now lets look at the number of sudden gains

```{r}
PCL_Data<-rawData %>% select(mrn,contains("pcl5_score")) %>% select(-contains("followup"))
cutoff<-define_crit1_cutoff(data_sd=rawData$pcl5_score_pastmonth_base,reliability=0.84)
vars<-colnames(PCL_Data)[2:8]
byperson <- create_byperson(data = rawData,
                    sg_crit1_cutoff=cutoff$reliable_change_value,
                    id_var_name = "mrn",
                    tx_start_var_name = "pcl5_score_pastmonth_base",
                    tx_end_var_name = "pcl5_score_post",
                    sg_var_list=vars,
                    identify_sg_1to2=T,
                    sg_measure_name = "PCL",
                    identify = "sg")
byperson <- mutate(byperson,sg_stable=if_else(!is.na(sg_reversal) & sg_reversal==0,1,0))
rawData<-byperson %>% select(mrn,sg_stable) %>%
  inner_join(rawData,by=c("mrn"))
chisq.test(rawData$sg_stable,rawData$followup_dropout)
table(rawData$sg_stable,rawData$followup_dropout)
```

looked at if there were significant differences between inidividuals who completed 3 month followup in both itps - there were a handful of signfiicant results, all of which failed to replicate between the two ITPs
